---
title: "personalized page ranks"
author: "Mitch Porter"
date: "2024-03-24"
output: html_document
---

```{r}
#the following code is directly from Keyes, et al, 2021 https://github.com/aike2863/food-web-and-ecosystem-service-robustness/tree/v1.0.0

```

```{r}
#load packages
library(tidyverse)
library(igraph)

#load data
nodes <- read.csv("nodes.csv")
links <- read.csv("links.csv")

#flip edges
links_flip <- links[,c(2,1,3)]
net <- graph.data.frame(links_flip, directed = T, vertices = nodes)

#run personalized pagerank for each individual ecosystem service
#FI

FI <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 1,0,0,0,0,0,0,0))$vector, vertextype=nodes$vertextype)
FI_SPP <- FI[!(FI$vertextype > 2),]
FI_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="3", links$Prey,NA))
FI_direct <- c(na.omit(FI_direct))
FI_support <- FI_SPP[!(FI_SPP$SpeciesID %in% FI_direct$SpeciesID),]
attach(FI_support)
FI_support <- FI_support[order(-prob),] #ordered from high to low prob
detach(FI_support)
write.csv(FI_support,"./Indirect_FI.csv")

#RM

RM <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,1,0,0,0,0,0,0))$vector, vertextype=nodes$vertextype)
RM_SPP <- RM[!(RM$vertextype > 2),]
RM_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="4", links$Prey,NA))
RM_direct <- c(na.omit(RM_direct))
RM_support <- RM_SPP[!(RM_SPP$SpeciesID %in% RM_direct$SpeciesID),]
attach(RM_support)
RM_support <- RM_support[order(-prob),] #ordered from high to low prob
detach(RM_support)
write.csv(RM_support,"./Indirect_RM.csv")

#HF

HF <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,1,0,0,0,0,0))$vector, vertextype=nodes$vertextype)
HF_SPP <- HF[!(HF$vertextype > 2),]
HF_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="5", links$Prey,NA))
HF_direct <- c(na.omit(HF_direct))
HF_support <- HF_SPP[!(HF_SPP$SpeciesID %in% HF_direct$SpeciesID),]
attach(HF_support)
HF_support <- HF_support[order(-prob),] #ordered from high to low prob
detach(HF_support)
write.csv(HF_support,"./Indirect_HF.csv")

#BP

BP <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,0,1,0,0,0,0))$vector, vertextype=nodes$vertextype)
BP_SPP <- BP[!(BP$vertextype > 2),]
BP_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="6", links$Prey,NA))
BP_direct <- c(na.omit(BP_direct))
BP_support <- BP_SPP[!(BP_SPP$SpeciesID %in% BP_direct$SpeciesID),]
attach(BP_support)
BP_support <- BP_support[order(-prob),] #ordered from high to low prob
detach(BP_support)
write.csv(BP_support,"./Indirect_BP.csv")

#CC

CC <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,0,0,1,0,0,0))$vector, vertextype=nodes$vertextype)
CC_SPP <- CC[!(CC$vertextype > 2),]
CC_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="7", links$Prey,NA))
CC_direct <- c(na.omit(CC_direct))
CC_support <- CC_SPP[!(CC_SPP$SpeciesID %in% CC_direct$SpeciesID),]
attach(CC_support)
CC_support <- CC_support[order(-prob),] #ordered from high to low prob
detach(CC_support)
write.csv(CC_support,"./Indirect_CC.csv")

#CS

CS <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,0,0,0,1,0,0))$vector, vertextype=nodes$vertextype)
CS_SPP <- CS[!(CS$vertextype > 2),]
CS_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="8", links$Prey,NA))
CS_direct <- c(na.omit(CS_direct))
CS_support <- CS_SPP[!(CS_SPP$SpeciesID %in% CS_direct$SpeciesID),]
attach(CS_support)
CS_support <- CS_support[order(-prob),] #ordered from high to low prob
detach(CS_support)
write.csv(CS_support,"./Indirect_CS.csv")

#CV

CV <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,0,0,0,1,0,0))$vector, vertextype=nodes$vertextype)
CV_SPP <- CV[!(CV$vertextype > 2),]
CV_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="9", links$Prey,NA))
CV_direct <- c(na.omit(CV_direct))
CV_support <- CV_SPP[!(CV_SPP$SpeciesID %in% CV_direct$SpeciesID),]
attach(CV_support)
CV_support <- CV_support[order(-prob),] #ordered from high to low prob
detach(CV_support)
write.csv(CV_support,"./Indirect_CV.csv")

#ED

ED <- data.frame(SpeciesID=nodes$Species, prob=page_rank(graph=net, damping = 0.85, directed = T, personalized = c(rep(0,times=nrow(nodes)-nrow(nodes[nodes$vertextype > 2,])), 0,0,0,0,0,0,0,1))$vector, vertextype=nodes$vertextype)
ED_SPP <- ED[!(ED$vertextype > 2),]
ED_direct <- data.frame(SpeciesID = ifelse(links$edgetype=="10", links$Prey,NA))
ED_direct <- c(na.omit(ED_direct))
ED_support <- ED_SPP[!(ED_SPP$SpeciesID %in% ED_direct$SpeciesID),]
attach(ED_support)
ED_support <- ED_support[order(-prob),] #ordered from high to low prob
detach(ED_support)
write.csv(ED_support,"./Indirect_ED.csv")

#calculating aggregate probability

all_support <- rbind(ED_support,CV_support,CS_support,CC_support,BP_support,HF_support,RM_support,FI_support)
all_support$SpeciesID <- as.factor(all_support$SpeciesID)
mean_support <- aggregate(prob ~ SpeciesID, all_support, mean) # calculate mean prob. for each sp.

attach(mean_support)
mean_support <- mean_support[order(-prob),]
detach(mean_support)

write.csv(mean_support, "importance.csv")